Who Cares About Crackdowns? Exploring Individual Donor Responses to Closing Civic Space Abroad

### 2020-06-07 manuscript/manuscript.Rmd
Finish most remaining TODOs

---
title: "Who Cares About Crackdowns? Exploring Individual Donor Responses to Closing Civic Space Abroad"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "June 5, 2020"
published: Working paper.
git-repo: https://example.com
abstract: >-
  TODO: UPDATE THIS: Over the past decade, international nonprofit organizations (INGOs) have faced increasing legal scrutiny and closed civic space abroad. While we know this has negatively impacted public funding flows, we know little about how this has affected private funding flows, especially be individuals. This turn to private aid introduces potentially new dynamics in INGO fundraising strategies, since individuals have a range of demographic characteristics and past experiences with philanthropy that shape their propensity to donate. Which demographic and experiential factors influence the decision to donate to legally besieged NGOs? We investigate this question using nationally representative results from a vignette-based experiment and a conjoint experiment conducted on individuals with a history of charitable-giving. Our findings have practical implications for INGOs that hope to raise funds from the general public amid closing civic space. 
reference-section-title: References
toc: false
mainfont: Lora
sansfont: IBM Plex Sans
---

@@

In Human Rights Watch's 2016 annual report, Executive Director Kenneth Roth argued that civil society was under more aggressive attack than at any time in recent memory. Over the past few decades, governments across the globe have been cracking down on civil society organizations (CSOs)—a phenomenon known as "closing civic space" in the policy arena [@CarothersBrechenmacher:2014; @CIVICUS:2017]. Consequently, official aid flows to NGOs have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While we know that there has been a negative impact on funding flows from public sector organizations like USAID, DFID, and others, we know relatively little about how this global crackdown affects private funding flows, in particular from individual donors.

@@

TODO ANDREW: [METHOD AND FINDINGS] We investigate these questions using a nationally representative survey of individuals in the U.S. with a previous history of charitable-giving. ONE LINE ABOUT EXPERIMENT. WHAT WE FIND

This study contributes to both to scholarship and practice in public policy, nonprofits and philanthropic-giving. Individual-level philanthropic giving abroad can help bridge socioeconomic disparities worldwide, especially given the stark inequality in global distribution of income. Knowledge of demographic traits and prior experiences that make certain individuals more likely to donate can be very important to nonprofits, as this information can help them craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Most CSOs and nongovernmental organizations (NGOs) working internationally have previously relied on government and foundation funding. But in the era of closing civic space, these groups may need to readjust their fundraising strategies to cater to private individuals.

@@

More broadly, while research in public policy and nonprofit studies has focused on charitable giving, philanthropy is not seen as relevant by international relations as a discipline [@youde2019]. However, philanthropy—particularly donations received by international non-governmental organizations (INGOs)—can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. As @youde2019 [p. 44] points out, it changes our conceptions of individuals being passive recipients to active entities that are not merely objects of governance, thus shedding light on the changing logic of global governance.

TODO ANDREW: [ROADMAP] Below, we first explain the different fundings for CSOs working internationally, examining trends in individual donor to philanthropy to such causes. We then take a look at the state of the existing literature on the key drivers of philanthropy by individuals. We describe our survey experiment. FINISH ROADMAP.

@@

Foreign aid to civil society, primarily via NGOs, has been steadily increasing over time. This is because many donors perceive that NGOs have superior ability to achieve intended outcomes. In countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013]. These groups provide services to large parts of the population in the Global South, especially where governments are unable to do so on a consistent basis. More broadly, a robust civil society is also considered to be one of the necessary preconditions for a healthy, functioning democracy due to its ability to generate social capital and hold governments accountable [@Putnam:1995].

@@

Thousands of international NGOs and CSOs in the Global South receive funds from a variety of public and private sources each year. Public funds (more commonly known as traditional donor aid) channels funds from official aid agencies through a variety of bilateral and multilateral institutions. Private funds, also newer international donors, included foundations, private corporations and individuals. In this paper, we specifically look at individual-level determinants to foreign giving. This is important to investigate as individual private donors may not necessarily mirror public donor choices. Surveys show that the American public is notoriously misinformed about US foreign aid [@norris:2017]. Further, foreign aid and democracy aid is often considered unresponsive to developments on the ground [@brainard2007; @Carothers:2015] and public aid decisions - especially on final recipients are not easily accessible. There may also be heterogeneity within how donor governments allocate funds. Emerging donor government differ from established donors in that they demand fewer human rights, governance or environmental conditions preferred by Western donors [@fenglerkharas].

With the advent of funding restrictions on NGOs globally and a consequent decline in official government aid and foundation funding, many groups are being forced to foreign individual donors. Giving to international affairs has been steadily increasing over the past two decades, with a noticeable increase over the past decade. Between 2014 and 2016, giving to international affairs increased by 14.1% [@givingusa2017report]. Most of this is driven by individuals making small-scale donations. In 2017, individuals gave \$286.7 billion - or approximately 70% of total giving to international affairs. This further increased to \$427.71 billion in 2018 - 85.7% of which was through individuals [givingusa:2019]. Given that a majority of this research is restricted to elite, high net-worth donors [@USTrust2014], these figures show that it is imperative to understand motivations and attributes of small-scale donors as well.

@@

A vast literature in philanthropic-giving looks at organizational-level factors that motivate individuals to donate [@BilodeauSlivinski:1997; @TBPrakash:2017]. However, the decision to donate to an NGO is not determined solely by an organization's characteristics. Donors have particular personality traits, preferences, and experiences that make them more or less likely to engage in philanthropy. A robust literature in nonprofit studies has looked at individual donor behavior and characteristics for giving to local-rather than international-causes. This literature has primarily focused on the importance of three main factors, especially when considering charitable behavior towards international causes: (1) social capital; (2) the role of higher education; (3) levels of religiosity; (4) wealth; and (5) national group attachment.

One major area of inquiry in philanthropic studies looks at the role of social capital, or the "networks, norms, and social trust that facilitate coordination and cooperation for mutual benefit" [@Putnam:1995, 67]. While quite broad, the literature further concentrates on a specific kind of social capital-associational capital, which refers to, "the networks of formal and informal relationships to which people are associated." [@havens2007, 240]. Prior research has found that those who participate in a variety of associations are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not), were both more likely to make charitable donations and make bigger donations. This is because volunteering exposes individuals to the need for donations to charities and nonprofits and made them feel more compelled to do so [@hossainlamb2012].

Higher education can also play an important role in philanthropic-giving decisions as such education fosters prosocial motivations. It also brings people into particular social networks that entail a higher level of solicitations [@BekkersWiepking:2011]. This in turn has been shown to be particularly vital in giving to international causes. @Micklewright:2009 show that higher education matters more for donating to international relief organizations than donating to domestic organizations. An additional year of education has been shown to increase the likelihood of giving to international causes by about 2 percentage points [@OktenOsili:2007].

Higher education can also play an important role in determining the kinds of causes that people donate to globally. A number of NGOs facing crackdowns abroad work on issues that may be seen as "contentious" - these include human rights, elections, corruption, advocacy, media freedom - causes that may have a harder time attracting funds due to their non-essential nature when compared to causes like health, education, sanitation, and social services. However, additional education has been shown to increase abstract thinking [@wiepkingmaas, 1978], which may be necessary to donate to causes that do not have quantifiable goals - which is particularly important for organizations with more "contentious" causes as part of their missions [@Bush:2015].

Levels of religiosity are also a significant predictor of the likelihood and level of giving. People who regularly attention religious services are more likely to have prosocial values, face a stronger norm of charitable-giving, and may feel pressure to conform that group and community standard [@wiepking2010]. Regular attendance may also strengthen individuals' beliefs in organizations' missions. However, religiosity does not necessarily have the same effect on giving to all causes. @neumayrhandy2019 find that religiosity is negatively associated with giving to environmental issues and animal welfare, while having an insignificant effect on giving to international relief.

@@

Overall, individuals with higher levels of income, education, and greater religious proclivities have been shown to not only be more likely to give internationally, but also to give higher amounts [@Rajan:2009; @CasaleBaumann:2015]. INSERT LINE ABOUT while we expect these determinants to hold similar significance in our survey, we do not specifically lay out hypotheses because this is exploratory/descriptive/we barely know about international giving so we don't have specific expectations about international giving during crackdown. BUT ALSO WHAT ARE WE DOING THAT HASN'T BEEN DONE YET.

TODO ANDREW: NEW GROUNDBREAKING STUFF = combining individual-level characteristics with organization characteristics - Looking at a persona is not typical - we often look at the effect of a single covariate, not a combination of them. But looking at the impact of a persona — which in all reality is a fabricated individual — has interesting appeal.

@@

We recruited survey respondents through Centiment, a commercial provider of high quality non-probability opt-in survey panels. Sentiment ensures panel quality by actively recruiting representative samples of the US population and rewards participants with monetary incentives. To participate in the study, respondents were first screened based on their charitable activities. Those who indicated that they gave to charity once every few years or never were disqualified. After screening, we collected a sample of 1,016 respondents, which is sufficient for model estimation.[^sample-size]

```{r child = "output/tables/tbl-sample-details.md"}

```

@tbl:sample-details provides a summary of the individual characteristics of survey respondents. Because we limited the sample to people who regularly make charitable contributions, most respondents have a favorable opinion of nonprofits and charities, with more than half reporting a 5 or greater level of trust in charities (on a 7-point scale). Moreover, nearly 90% indicated that it is important to trust charities. Additionally, the majority of respondents volunteer at least once a year, donate to charity at least once a year, and give between \$1–\$999 annually. Most rarely attend church, but feel that religion is important. Respondents are also engaged civically (three quarters voted in the last election) and follow national news closely (70% follow the news daily). Respondents are fairly balanced ideologically, with the majority reporting a 3, 4, or 5 on a 7-point scale of liberal–conservative political ideology.

@@

In this paper, we focus instead on exploring which kinds of donors are more or less likely to consider an NGO's issue area, funding sources, and host country politics when deciding to donate. Disentangling the effect of every individual level-characteristics on every combination of organizational characteristics is unwieldy and near impossible. Given the organizational attributes presented in our experiment, and the possible individual characteristics that respondents could select, there are billions of possible combinations and iterations of organizational and individual attributes.

Instead of trying to work through all these countless combinations—most of which rarely occur in the actual population of potential donors—we rely on simulation methods to examine the effect of more general combinations of individual characteristics. We use market segmentation methods to construct simulated personas that interact with simulated organizations in a market for charitable contributions. To simplify the many possible combinations of individual characteristics, we generate 32 distinct personas with attributes that vary along different dimensions of demographics, politics and public affairs experience, and social attitudes (see Table XXX). We also generate 24 simulated organizations that vary by issue area, relationship with host government, and source of funding. This combination of personas and organizations yields 768 hypothetical scenarios where potential donors can interact with different organizations.

TODO ANDREW: Simplified persona table + organization table

We then combine these different scenarios of donor and organization attributes with the coefficients from a multinomial multilevel regression model to calculate the probability that each simulated persona has of giving to specific combinations of organizational attributes. Any individual characteristics that we do not explicitly vary in Table X are held at their means or modal values from the survey sample. Table X provides an example of the simulation output, showing how two different personas are predicted to allocate donations to organizations with various characteristics. The proportion of donations from each persona sums to 100%, as every persona is guaranteed to make a donation.

TODO ANDREW: Two simplified sample personas

@@

Figures @fig:issue-social, @fig:relationship-social, and @fig:funding-social show the results of the market simulation across these three dimensions of persona characteristics, varying public affairs knowledge, political ideology, and social trust. Personas respond differently to organizational issue areas. In @fig:issue-social, those who follow the news and travel abroad are more likely to prefer donating to human rights organizations than their less knowledgable and less travelled counterparts. Personas who do not follow the news or travel show a greater preference for emergency response and refugee relief organization, similar to the income effects seen previously. There are no strong ideological effects across issue areas. Both conservative and liberal personas tend to follow similar patterns of preference across NGO issue. Social trust plays a substantial role in determining simulated market share, however. Those who trust social and charitable institutions are very likely to prefer donating to human rights organizations, regardless of their public affairs knowledge or political leanings. In all circumstances, those with high levels of trust have between a 7–10% chance of donating to a human rights organization, on average. Personas with low levels of trust, on the other hand, strongly avoid human rights organizations and prefer instead to donate to refugee and emergency response NGOs. The difference is substantial—a liberal persona with high levels of public affairs knowledge and experience will give to a human rights organization 10% of the time, on average if they have high levels of trust, but less than 3% of the time if they are less trusting. Trust in charities and institutions is a key factor in deciding to donate to more contentious human rights organizations.

@@

Figure @fig:relationship-social shows how personas respond to different NGO–host government relationships. Public affairs knowledge and experience has little effect on donor preferences—the average market shares are roughly identical across the two columns of the figure. Similarly, there are no ideological differences across the rows of the figure. Persona preferences toward friendly, criticized, and legally suppressed organizations do not change much across different ideological views. The strongest, most overriding factor in determining donor preferences when considering an organization's relationship with its host country is social trust. Personas with high levels of social trust are overwhelmingly most likely to donate to organizations with friendly relationships with their host governments, while those with low levels of trust avoid friendly NGOs almost entirely.

@@

With the turn towards private philanthropy in the face of the global crackdown on civil society and declining official aid and official development assistance, INGOs need to be aware of the dynamics of individual donor preferences. WHAT WE FIND

@@

Finally, it is important to acknowledge the role played by civil society organizations in dealing with health, poverty, and disaster - social needs that are tremendously important during the COVID-19 pandemic. This survey was conducted in July (?) 2019 - can we expect insights from this survey to hold even during this time period, even as the world sees an economic recession? We argue that this is the case for two main reasons. First, prior research demonstrates that altruism and prosocial behavior are more common during and after disasters (cite?). Second, the most recent instance of the 2008 recession did not uniformly impact all demographic groups in the U.S. - individuals with lower levels of education, income, or wealth showed a statistically significant decline in charitable-giving. However, philanthropic patterns were not significantly different for other income brackets.

>>>>

---
title: "Who Cares About Crackdowns? Exploring Individual Donor Responses to Closing Civic Space Abroad"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "June 5, 2020"
published: Working paper.
git-repo: https://example.com
abstract: >-
  Over the past decade, international nonprofit organizations (INGOs) have faced increasing legal scrutiny and closed civic space abroad. While we know this has negatively impacted public funding flows, we know little about how this has affected private funding flows, especially be individuals. This turn to private aid introduces potentially new dynamics in INGO fundraising strategies, since individuals have a range of demographic characteristics and past experiences with philanthropy that shape their propensity to donate. Which demographic and experiential factors influence the decision to donate to legally besieged NGOs? We investigate this question using a market simulation based on a nationally representative conjoint survey experiment conducted on individuals with a history of charitable giving. We find that trust in political institutions and charities is enormously influential in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to human rights INGOs with friendly relationships with host governments, but that support evaporates if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do face criticism or crackdown abroad. Our findings have practical implications for INGOs that hope to raise funds from the general public amid closing civic space. 
reference-section-title: References
toc: false
mainfont: Lora
sansfont: IBM Plex Sans
---

@@

In Human Rights Watch's 2016 annual report, Executive Director Kenneth Roth argued that civil society was under more aggressive attack than at any time in recent memory. Over the past few decades, governments across the globe have been cracking down on civil society organizations (CSOs)—a phenomenon known as "closing civic space" in the policy arena [@CarothersBrechenmacher:2014; @CIVICUS:2017]. Consequently, official aid flows to nongovernmental organizations (NGOs) have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While we know that there has been a negative impact on funding flows from public sector organizations like USAID, DFID, and others, we know relatively little about how this global crackdown affects private funding flows, in particular from individual donors.

@@

We explore these questions using a simulated market for philanthropic donations, based on data and models from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing research to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that specific individual characteristics help shape donor preferences for donating to international NGOs, and that an INGO's target issue area and its relationship with host government interact strongly with these characteristics. In particular, personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism tend to prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with friendly relationships with host governments and prefer instead to donate to NGOs facing more restrictions and working on less contentious issues like emergency response and refugee relief. This may indicate that donors with high trust and support for NGOs could be fair-weather friends that drop their support when their preferred organizations meet resistance abroad.

This study contributes to both to scholarship and practice in public policy, nonprofits and charitable giving. Individual-level philanthropic giving abroad can help bridge socioeconomic disparities worldwide, especially given the stark inequality in global distribution of income. Knowledge of demographic traits and prior experiences that make certain individuals more likely to donate can be very important to nonprofits, as this information can help them craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Most CSOs and NGOs working internationally have previously relied on government and foundation funding. But in the era of closing civic space, these groups may need to readjust their fundraising strategies to cater to private individuals.

@@

More broadly, while research in public policy and nonprofit studies has focused on charitable giving, philanthropy is not seen as relevant by international relations as a discipline [@youde2019]. However, philanthropy—particularly donations received by international nongovernmental organizations (INGOs)—can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. As @youde2019 [p. 44] points out, it changes our conceptions of individuals being passive recipients to active entities that are not merely objects of governance, thus shedding light on the changing logic of global governance.

Below, we first explore how the global crackdown on civil society has led many CSOs working internationally to shift their fundraising strategies towards individual private donors rather than more restricted government and foundation sources. We then examine the state of the existing literature on the key drivers of individual charitable giving, as understanding what motivates private donations is crucial to more effective CSO fundraising. We then describe our survey experiment and simulation methods and explore the results. We conclude with a discussion of the implications our findings have on general INGO funding in the age of constrained civil society space.

@@

Foreign aid to civil society—and NGOs in particular—has steadily increased over time. One primary reason for this turn to civil society is that many government and foundation donors perceive that NGOs have a superior ability to achieve intended outcomes. In countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013]. These groups provide services to large parts of the population in the Global South, especially where governments are unable to do so on a consistent basis. More broadly, a robust civil society is also considered to be one of the necessary preconditions for a healthy, functioning democracy due to its ability to generate social capital and hold governments accountable [@Putnam:1995].

@@

Thousands of international NGOs and CSOs in the Global South receive funds from a variety of public and private sources each year. Public funds (more commonly known as traditional donor aid) channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. In this paper, we are most interested in individuals as a funding source, and we examine individual-level determinants to foreign giving. This is important to investigate as individual private donors may not necessarily mirror public donor choices. Surveys show that the American public is notoriously misinformed about US foreign aid [@norris:2017]. Further, foreign aid and democracy aid is often considered unresponsive to developments on the ground [@brainard2007; @Carothers:2015] and public aid decisions—especially on final recipients— are not easily accessible. There is also substantial variety in how donor governments allocate funds. Emerging donor governments differ from established donors in that they demand fewer human rights, governance or environmental conditions preferred by Western donors [@fenglerkharas].

With the advent of funding restrictions on NGOs globally and a subsequent decline in official government aid and foundation funding, many organizations are being forced to individual donors abroad. Donations to international affairs organizations has steadily increased over the past two decades, with a noticeable rise in the past decade. Between 2014 and 2016, giving to international affairs increased by 14.1% [@givingusa2017report]. Most of this is driven by individuals making small-scale donations. In 2017, individuals gave \$286.7 billion, or approximately 70% of total giving to organizations working in international affairs. This further increased to \$427.71 billion in 2018, 85.7% of which was through individuals [givingusa:2019]. Given that a majority of this research is restricted to elite, high net-worth donors [@USTrust2014], these figures show that it is imperative to understand motivations and attributes of small-scale donors as well.

@@

A vast literature in charitable giving looks at organizational-level factors that motivate individuals to donate [@BilodeauSlivinski:1997; @TBPrakash:2017]. However, the decision to donate to an NGO is not determined solely by an organization's characteristics. Donors have particular personality traits, preferences, and experiences that make them more or less likely to engage in philanthropy. A robust literature in nonprofit studies has looked at individual donor behavior and characteristics for giving to local—rather than international—causes. This literature has primarily focused on the importance of three main factors, especially when considering charitable behavior towards international causes: (1) social capital; (2) the role of higher education; (3) levels of religiosity; (4) wealth; and (5) national group attachment.

One major area of inquiry in philanthropic studies examines the role of social capital, or the "networks, norms, and social trust that facilitate coordination and cooperation for mutual benefit" [@Putnam:1995, 67]. While quite broad, the literature further concentrates on a specific kind of social capital—associational capital, or "the networks of formal and informal relationships to which people are associated." [@havens2007, 240]. Prior research has found that those who participate in a variety of associations are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not) were both more likely to make charitable donations and make bigger donations. This is because volunteering exposes individuals to the need for donations to charities and nonprofits and made them feel more compelled to do so [@hossainlamb2012].

Higher education can also play an important role in charitable giving decisions as such education fosters prosocial motivations. It also brings people into particular social networks that entail a higher level of solicitations [@BekkersWiepking:2011]. This in turn has been shown to be particularly vital in giving to international causes. @Micklewright:2009 show that higher education matters more for donating to international relief organizations than donating to domestic organizations. An additional year of education has been shown to increase the likelihood of giving to international causes by about 2 percentage points [@OktenOsili:2007].

Higher education also helps determine the kinds of causes that people donate to globally. A number of NGOs facing crackdowns abroad work on issues that may be seen as "contentious," including human rights, elections, corruption, advocacy, media freedom—all issues that may have a harder time attracting funds due to their non-essential nature when compared to causes like health, education, sanitation, and social services. However, additional education has been shown to increase abstract thinking [@wiepkingmaas, 1978], which may be necessary to donate to causes that do not have quantifiable goals, which is particularly important for organizations with more "contentious" causes as part of their missions [@Bush:2015].

Levels of religiosity are also a significant predictor of the likelihood and level of giving. People who regularly attention religious services are more likely to have prosocial values, face a stronger norm of charitable giving, and may feel pressure to conform that group and community standard [@wiepking2010]. Regular attendance may also strengthen individuals' beliefs in organizations' missions. However, religiosity does not necessarily have the same effect on giving to all causes. @neumayrhandy2019 find that religiosity is negatively associated with giving to environmental issues and animal welfare, while having an insignificant effect on giving to international relief.

@@

Overall, individuals with higher levels of income, education, and greater religious proclivities have been shown to not only be more likely to give internationally, but also to give higher amounts [@Rajan:2009; @CasaleBaumann:2015]. While we expect these determinants to be reflected in our survey experiment, we do not lay out specific hypotheses. Because there is so little research about international donations during legal crackdown by host governments, our analysis and results are exploratory and descriptive. Our innovative experimental approach, however, provides rich quantitative details of individual and organizational characteristics simultaneously and reveals unique insights into donor motivations that both confirm and expand previous research.

@@

In June 2019, We recruited survey respondents through Centiment, a commercial provider of high quality non-probability opt-in survey panels. Sentiment ensures panel quality by actively recruiting representative samples of the US population and rewards participants with monetary incentives. To participate in the study, respondents were first screened based on their charitable activities. Those who indicated that they gave to charity once every few years or never were disqualified. After screening, we collected a sample of 1,016 respondents, which is sufficient for model estimation.[^sample-size]

\newpage

```{r child = "output/tables/tbl-sample-details.md"}
```

@tbl:sample-details provides a summary of the individual characteristics of survey respondents. Because we limited the sample to people who regularly make charitable contributions, most respondents have a favorable opinion of nonprofits and charities, with more than half reporting a 5 or greater level of trust in charities (on a 7-point scale). Moreover, nearly 90% indicated that it is important to trust charities. Additionally, the majority of respondents volunteer at least once a year, donate to charity at least once a year, and give between \$1–\$999 annually. Most rarely attend church, but feel that religion is important. Respondents are also civically engaged (three quarters voted in the last election) and follow national news closely (70% follow the news daily). Respondents are fairly balanced ideologically, with the majority reporting a 3, 4, or 5 on a 7-point scale of liberal–conservative political ideology.

@@

In this paper, we focus instead on exploring which kinds of donors are more or less likely to consider an NGO's issue area, funding sources, and host country politics when deciding to donate. Disentangling the effect of every individual level-characteristics on every combination of organizational characteristics is unwieldy and near impossible. Given the organizational attributes presented in our experiment, and the possible individual characteristics that respondents could select, there are billions of possible combinations and iterations of organizational and individual attributes—many of which rarely occur in the actual population of potential donors. In our previous work—and in most research that relies on regression model—the analytic focus is on individual regression coefficients and the effect of single covariates of interest. To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates at simultaneously

We use market segmentation methods to construct simulated personas that interact with simulated organizations in a market for charitable contributions. To simplify the many possible combinations of individual characteristics, we generate 32 distinct personas with attributes that vary along different dimensions of demographics, politics and public affairs experience, and social attitudes (see @tbl:persona-attributes). We also generate 24 simulated organizations that vary by issue area, relationship with host government, and source of funding (see @tbl:organization-attributes). This combination of personas and organizations yields 768 hypothetical scenarios where potential donors can interact with different organizations.

```{r child = "output/tables/tbl-persona-attributes.md"}
```

```{r child = "output/tables/tbl-organization-attributes.md"}
```

We then combine these different scenarios of donor and organization attributes with the coefficients from a multinomial multilevel regression model to calculate the probability that each simulated persona has of giving to specific combinations of organizational attributes. Any individual characteristics that we do not explicitly vary in  @tbl:persona-attributes are held at their means or modal values from the survey sample. @tbl:sim-output provides an example of the simulation output, showing how two different personas are predicted to allocate donations to organizations with various characteristics. The proportion of donations from each persona sums to 100%, as every persona is guaranteed to make a donation.

```{r child = "output/tables/tbl-sim-output.md"}
```

@@

Figures @fig:issue-social, @fig:relationship-social, and @fig:funding-social show the results of the market simulation across these three dimensions of persona characteristics, varying public affairs knowledge, political ideology, and social trust. Personas respond differently to organizational issue areas. In @fig:issue-social, those who follow the news and travel abroad are more likely to prefer donating to human rights organizations than their less knowledgeable and less traveled counterparts. Personas who do not follow the news or travel show a greater preference for emergency response and refugee relief organization, similar to the income effects seen previously. There are no strong ideological effects across issue areas. Both conservative and liberal personas tend to follow similar patterns of preference across NGO issue. Social trust plays a substantial role in determining simulated market share, however. Those who trust social and charitable institutions are very likely to prefer donating to human rights organizations, regardless of their public affairs knowledge or political leanings. In all circumstances, those with high levels of trust have between a 7–10% chance of donating to a human rights organization, on average. Personas with low levels of trust, on the other hand, strongly avoid human rights organizations and prefer instead to donate to refugee and emergency response NGOs. The difference is substantial—a liberal persona with high levels of public affairs knowledge and experience will give to a human rights organization 10% of the time, on average if they have high levels of trust, but less than 3% of the time if they are less trusting. Trust in charities and institutions is a key factor in deciding to donate to more contentious human rights organizations.

@@

Figure @fig:relationship-social shows how personas respond to different NGO–host government relationships. Public affairs knowledge and experience has little effect on donor preferences, as the average market shares are roughly identical across the two columns of the figure. Similarly, there are no ideological differences across the rows of the figure. Persona preferences toward friendly, criticized, and legally suppressed organizations do not change much across different ideological views. The strongest, most overriding factor in determining donor preferences when considering an organization's relationship with its host country is social trust. Personas with high levels of social trust are overwhelmingly most likely to donate to organizations with friendly relationships with their host governments, while those with low levels of trust avoid friendly NGOs almost entirely.

@@

With the turn towards private philanthropy in the face of the global crackdown on civil society and declining official aid and official development assistance, INGOs need to be aware of the dynamics of individual donor preferences. Based on our simulated market of 24 organizations and 32 persona profiles, we find several important trends. On average, human rights organizations tend to attract personas who are wealthier, follow national and international news, and travel abroad, while donors who are less wealthy, follow the news less often, and do not travel are more attracted to organizations working on  less contentious like emergency response and refugee relief. More importantly, we found that trust in political institutions and in charitable organizations in general matters substantially in shaping donor preferences. Personas with high levels of social and charitable trust prefer donating to human rights organizations that have friendly relationships with their host governments, while personas with low trust are drawn more to NGOs emergency response and refugee relief organizations with unfriendly relationships with their host governments. Personas with low social trust rarely prefer to donate to any type of organization with friendly host-country relationships, while personas with high social trust tend to avoid organizations with unfriendly relationships. This could imply that donors that are the most trusting and involved might be fair-weather friends that drop their support when charities face government criticism and crackdown.

TODO: All this conclusionary stuff + so what/implications?

@@

Finally, it is important to acknowledge the role played by civil society organizations in dealing with health, poverty, and disaster - social needs that are tremendously important during the COVID-19 pandemic. This survey was conducted in June 2019 - can we expect insights from this survey to hold even during this time period, even as the world sees an economic recession? We argue that this is the case for two main reasons. First, prior research demonstrates that altruism and prosocial behavior are more common during and after disasters (cite?). Second, the most recent instance of the 2008 recession did not uniformly impact all demographic groups in the U.S. - individuals with lower levels of education, income, or wealth showed a statistically significant decline in charitable-giving. However, philanthropic patterns were not significantly different for other income brackets.
